Finding the True MVP

### 2024-12-14 paper/paper.qmd
Small edits to spelling and grammar

We expect a positive relationship between MVP Index and points, assists, rebounds, blocks, steals, and win shares, as these variables collectively define the MVP Index. Additionally, we anticipate a positive relationship because all these statistics indicate strong player performance. The larger any one of these statistics is, the larger the MVP Index will be for the respective player. However, a player who does not acquire many blocks or assists will not necessarily have a low MVP rating; they will simply need to compensate with higher performance in other areas. As shown in @sec-model-details, the model has a p-value of less than $2.2 \times 10^{-16}$, which is very low and gives us confidence that the model is accurate, as well as an $R^{2}$ of $R^{2} = 0.8992$, with a corresponding $R = 0.9483$.

@@

In contrast, players who showcase moments of sheer brilliance often become the face of the league. These athletes might not exhibit the same game to game reliability, but their ability to dominate in spectacular fashion leaves a lasting impression on fans and voters alike. Peak performers can change the momentum of a season with game-winning shots, record-breaking performances, or memorable battles against top opponents. However, their inconsistency can leave teams vulnerable during less-publicized matchups or critical stretches of the season.Ultimately, the MVP conversation remains an evolving debate, shaped by both quantitative measures and subjective interpretations of what makes a player "most valuable."

@@

Next, team records are also incredibly important for MVP, if a player played incredibly one season but barely made the playoffs, or even did not at all, the player likely should not be receiving the MVP award, as how much worse could said team get if they weren't that good with the so called "best" player in the league. Unfortunately we were notable to inclue team record, team win percentage, or even player win percentage if we do not want to include the games a player may have missed. Unfortunately this data was not included in either type of data set that we sourced from [@bbref], and due to time constraints downloading potentially dozens if not hundreds additional data sets did not seem to be the best use of time in our case.

@@

Our next steps, would like be to monitor this NBA season closely potentially add some of the data mentioned earlier in @sec-weak, such as win percentage, and individual historical statistics to more accurate predict the model. We would also like to revisit this paper tat the end of this NBA season once the NBA awards are announced and see how we could improve our model and how accurate it was in predicting the MVP Index of the next MVP.

>>>>

We expect a positive relationship between MVP Index and points, assists, rebounds, blocks, steals, and win shares, as these variables collectively define the MVP Index. Additionally, we anticipate a positive relationship because all these statistics indicate strong player performance. The larger any one of these statistics is, the larger the MVP Index will be for the respective player. However, a player who does not acquire many blocks or assists will not necessarily have a low MVP rating; they will simply need to compensate with higher performance in other areas. As shown in @sec-model-details, the model has a p-value of less than $2.2 \times 10^{-16}$, which is very low and gives us confidence that the model is accurate, as well as an $R^{2}$ score of $R^{2} = 0.8992$, with a corresponding $R = 0.9483$.

@@

In contrast, players who showcase moments of sheer brilliance often become the face of the league. These athletes might not exhibit the same game to game reliability, but their ability to dominate in spectacular fashion leaves a lasting impression on fans and voters alike. Peak performers can change the momentum of a season with game-winning shots, record-breaking performances, or memorable battles against top opponents. However, their inconsistency can leave teams vulnerable during less-publicized match ups or critical stretches of the season.Ultimately, the MVP conversation remains an evolving debate, shaped by both quantitative measures and subjective interpretations of what makes a player "most valuable."

@@

Next, team records are also incredibly important for MVP, if a player played incredibly one season but barely made the playoffs, or even did not at all, the player likely should not be receiving the MVP award, as how much worse could said team get if they weren't that good with the so called "best" player in the league. Unfortunately we were notable to include team record, team win percentage, or even player win percentage if we do not want to include the games a player may have missed. Unfortunately this data was not included in either type of data set that we sourced from [@bbref], and due to time constraints downloading potentially dozens if not hundreds additional data sets did not seem to be the best use of time in our case.

@@

Ideally, our next steps, would be to monitor this NBA season closely potentially add some of the data mentioned earlier in @sec-weak, such as win percentage, and individual historical statistics to more accurately predict the model. We would also like to revisit this research at the end of this NBA season once the NBA awards are announced and see how we could improve our model and how accurate it was in predicting the MVP Index of the next MVP.
